---
title: "Assignment 1"
---
Ps. All the codes below has written with the combined power of AI & HI.
If you want to reach promts of codes that is used while preparing assignment, you can access the pdf document. 
📄 [Download Prompts](prompts.pdf){.download target="_blank"}



My first assignment has three parts.

## (a)

```{r}
1 + 1
```

## (b.1)

```{r}
compute_stats <- function(data) {
  results <- list()
  
  for (col_name in colnames(data)) {
    column_values <- data[[col_name]]
    
    stats <- list(
      mean = mean(column_values, na.rm = TRUE),
      median = median(column_values, na.rm = TRUE),
      variance = var(column_values, na.rm = TRUE),
      IQR = IQR(column_values, na.rm = TRUE),
      min = min(column_values, na.rm = TRUE),
      max = max(column_values, na.rm = TRUE)
    )
    
    results[[col_name]] <- stats
  }
  
  return(results)
}

data(mtcars)

stats_results <- compute_stats(mtcars)

print(stats_results)

```

## (b.2)

```{r}
compute_stats_sapply <- function(data) {
  stats <- sapply(data, function(column) {
    list(
      mean = mean(column, na.rm = TRUE),
      median = median(column, na.rm = TRUE),
      variance = var(column, na.rm = TRUE),
      IQR = IQR(column, na.rm = TRUE),
      min = min(column, na.rm = TRUE),
      max = max(column, na.rm = TRUE)
    )
  }, simplify = FALSE)  
  
  return(stats)
}

data(mtcars)

stats_results_sapply <- compute_stats_sapply(mtcars)

print(stats_results_sapply)

```

## (b.3)

```{r}
compute_stats_apply <- function(data) {
  stats_matrix <- apply(data, 2, function(column) {
    c(
      mean = mean(column, na.rm = TRUE),
      median = median(column, na.rm = TRUE),
      variance = var(column, na.rm = TRUE),
      IQR = IQR(column, na.rm = TRUE),
      min = min(column, na.rm = TRUE),
      max = max(column, na.rm = TRUE)
    )
  })
  
  return(as.data.frame(stats_matrix))  
}

stats_results_apply <- compute_stats_apply(as.matrix(mtcars))

print(stats_results_apply)

```

## (c)

```{r}
library(dslabs)
data("na_example")
na_example
```

```{r}

na_count <- sum(is.na(na_example))

print("Here, you can see the number of NA in this dataset:")
print(na_count)

```

```{r}

na_indexes <- which(is.na(na_example))

print("Here, you can basicly see the indexes of NA data:")
print(na_indexes)

```

```{r}

library(dslabs)
data("na_example")

mean_value <- mean(na_example, na.rm = TRUE)
sd_value <- sd(na_example, na.rm = TRUE)

print(paste("Mean:", mean_value))
print(paste("Standard Deviation:", sd_value))


```

```{r}

library(dslabs)
data("na_example")

mape <- function(actual, predicted) {
  return(mean(abs((actual - predicted) / actual)) * 100)
}

library(dslabs)
data("na_example")

median_value <- median(na_example, na.rm = TRUE)
version1 <- na_example
version1[is.na(version1)] <- median_value

set.seed(42)
non_na_values <- na_example[!is.na(na_example)]
version2 <- na_example
version2[is.na(version2)] <- sample(non_na_values, sum(is.na(version2)), replace = TRUE)

means <- c(mean(na_example, na.rm = TRUE),
           mean(version1, na.rm = TRUE),
           mean(version2, na.rm = TRUE))

sds <- c(sd(na_example, na.rm = TRUE),
         sd(version1, na.rm = TRUE),
         sd(version2, na.rm = TRUE))

mape_means_v1 <- mape(means[1], means[2])
mape_means_v2 <- mape(means[1], means[3])
mape_sds_v1 <- mape(sds[1], sds[2])
mape_sds_v2 <- mape(sds[1], sds[3])

stats_df <- data.frame(
  Version = c("Original", "Median Imputed", "Random Imputed"),
  Mean = means,
  Standard_Deviation = sds,
  MAPE_Mean = c(NA, mape_means_v1, mape_means_v2),  # MAPE for mean
  MAPE_SD = c(NA, mape_sds_v1, mape_sds_v2)         # MAPE for standard deviation
)

print(stats_df)




```

As seen on the table the version 2 array's mean and SD values are closer to the original dataset. It means that, to obtain a new dataset with the characteristic of original one, changing NA values with the randomly selected data from original dataset is more relevant.
